Supplementary material

### 2025-06-27 code/supplementary.Rmd
updated supplementary material and manuscript

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

@@

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      # cache = T,
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/manuscript/figs/")

@@

# Introduction

## Conventional M&V

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. Specifically, an M&V
analyst should account for two types of baseline changes: (1) static
changes, which are known building operational changes: such as
renovation, equipment addition or removal; and (2) gradual changes,
which often occur subtly and unknown, such as the incremental adoption
of LED lighting or more efficient plug loads. Identifying and
quantifying these changes are particularly labor-intensive and
challenging. Consequently, those limitations reduce the feasibility of
M&V and complicates the quantification of estimation uncertainty, thus
impacting the accuracy and timeliness of savings assessments.

## Randomized M&V

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance on two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defined stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for 1 intervention, the balanced
randomized schedule would equally sample 5 Mondays with the baseline
strategy in operation and 5 Mondays with the intervention strategy in
operation. The limitation of this method is that it is only applicable
to a subset of retrofit projects where interventions can be easily
switched on and off. However, for all applicable use cases, it allows
analysts to detect energy savings sequentially shortly after the test
begins. In addition, once the desired estimate uncertainty target is
achieved, analysts can terminate the M&V study and switch to 100%
intervention. The key advantage of randomization is that if control
strategies are sampled with equal probability, the influence of other
confounding factors such as occupancy change (commonly known as one of
the non-routine events) and other more subtle long term changes in
building energy use are likely to be evenly distributed among
measurements, leading to a more accurate and unbiased assessment of the
intervention effect.

## BDG2 dataset

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
running the M&V methods.

## Literature review

### Whole building approach

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
whole-building approach focuses on the accuracy of baseline modeling,
exploring model performance from simple regression models to more
complex machine learning techniques. One study reviewed various models
suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

### Changes in building baseline measurements

Changes in a building that influences its energy usage are commonly
refered to as non-routine events in M&V and are typically unrelated to
the intervention strategy. Current guidelines define non-routine events
as 'static factors' that needed to be adjusted after projecting baseline
in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine events as it also
represents a change in facility operation but in a longer term.

## Objectives

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

1.  Determine, using a large sample of buildings, how much more
    accurately the randomized M&V method would estimate the savings of
    the proposed intervention compared to the conventional M&V, and how
    much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset. The BDG2 dataset
    contains real-world building energy usage data collected over two
    years, inherently contains varying degrees of usage changes. By
    virtually introducing intervention effects into the existing
    dataset, we can calculate the ground-truth energy savings, which
    serve as a reference point for method comparison.

3.  Open-source implementation of the proposed randomized M&V method. We
    ensured the reproducibility of the method by making the analysis
    code open source including randomized schedule generation,
    sequential statistical analysis, energy modeling and normalized
    saving calculation. In addition, we also included code for extended
    use cases of the randomized method, such as changing sampling ratio
    and sampling intervals. Using the available open resources, building
    analysts should be able to seamlessly integrate and apply them in
    their own M&V projects.

# Method

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

knitr::include_graphics(paste0(fig_path, "flowchart.png"))
```

## Filter and clean dataset

We extracted the electricity measurements from the BDG2 dataset and
filter out all qualified buildings based on the following criteria:

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

2.  Mean electricity usage \> 0 kWh: target buildings should have active
    electricity usage.

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage in the second year should be
    less than 25% of that in the first year.

4.  Electric EUI \< 750 kWh/$m^2$: excludes buildings at the top 5% of
    electrical energy usage intensity according to the statistics
    provided by the Building Performance Database
    [@lawrenceberkeleynationallab; @mathew_big-data_2015].

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water set point reset
    control.

```{r stablesite, fig.cap = "Site summary of the filtered buildings from BDG2 dataset (counts < 10 are omitted for visualization; left: aggregated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

set3 <- colorRampPalette(brewer.pal('Set3',n=12))
type_colors <- setNames(set3(13), all_types_variable$type)

stable_set <- df_energy_stable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p1 <- variable_set %>% 
  left_join(stable_set, by = "type") %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 10, n, "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

stable_set <- df_energy_stable %>% 
  left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(stable = n()) %>%
  ungroup()

variable_set <- df_energy_variable %>% 
   left_join(df_loc, by = "site") %>% 
  group_by(type, location) %>%
  distinct(name) %>%
  summarise(variable = n()) %>%
  ungroup()

p2 <- variable_set %>% 
  left_join(stable_set, by = c("type", "location")) %>% 
  mutate(stable = replace_na(stable, 0)) %>% 
  group_by(location, type) %>% 
  summarise(n = stable + variable) %>% 
  ungroup() %>% 
  group_by(location) %>%
  mutate(proportion = n / sum(n),
         total = sum(n),
         ymax = cumsum(proportion),
         ymin = c(0, head(ymax, n = -1))) %>%
  mutate(label_pos = (ymax + ymin) / 2) %>%
  ungroup() %>%
  group_by(type) %>%
  mutate(order = sum(n)) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, order, .desc = F)) %>%
  ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3.2, fill = type)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  facet_wrap(~ location, nrow = 3) +
  labs(x = NULL,
       y = NULL,
       subtitle = "For each location",
       fill = NULL) +
  scale_fill_manual(values = type_colors) +
  geom_text(aes(x = 3.7, y = label_pos, label = ifelse(n > 10, n, ""))) +
  geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black", check_overlap = T) +
  theme(legend.direction = "horizontal",
        axis.text = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

ggarrange(p1, p2,
          ncol = 2, nrow = 1,
          widths = c(0.2, 1),
          common.legend = TRUE,
          legend = "bottom") +
  plot_annotation(title = "Summary of filtered buildings and types")
```

For all qualified buildings, we further added one more stringent
criterion to explore all buildings with very stable electricity usage
between the two years:

1.  No statistically significant difference (P-value \> 0.05) between
    the two-year electricity usage, which are sometimes assumed to be
    the case for whole-building measurement and verification.

As a result, only 66 buildings are labeled in this 'stable' subset (out
of total of 600 buildings) implying the assumption that buildings are
typically stable over the timeframe involved in conventional M&V is
rarely true. As mentioned in Section \@ref(non-routine-events-impact),
such variability is largely associated with non-routine events,
especially operational drifts since the assessment, even under ideal
situations, last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that resets the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller operation\nbelow 10 °C") +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

$$
\text{perc_savings = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

|                        |                                                                  |       |
|-------------------|----------------------------------|-------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

## Run M&V methods

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post- comparison on a 12-month baseline and
12-month intervention timeline. The randomized method defines sampling
requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

-   Block by day of the week with a block period of 12 weeks.

and stopping criteria as:

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of the satisfied blocking
    period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data sampled by both strategies.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
    and calculated SPRT statistics either falls below the lower
    threshold or exceeds the upper threshold.

-   Sample strategies at an equal sampling ratio (50% baseline, 50%
    intervention).

# Results

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step representing 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top for reference.

```{r abstract, fig.cap="Overall comparison results between the conventional M&V method and the proposed randomized M&V method (both upon reaching the stopping criteria for each case, and over a fixed two-year period).", fig.height=11, fig.width=12}

#### GA ####
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

@@

# plot for the variable subset
rand_eob_V <- df_seq_fs_variable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site"))%>% 
  mutate(diff = abs(savings - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

# TMY
# plot for the stable subset
rand_eob_S <- df_sprt_all_stable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_tmy_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_S <- rbind(rand_eob_S, conv_S, rand_final_S)

# plot for the variable subset
rand_eob_V <- df_sprt_all_variable %>% 
  filter(seq == "eob") %>% 
  select(-c(seq, n_weeks)) %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - annual), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_tmy_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = abs(savings - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

df_TW_V <- rbind(rand_eob_V, conv_V, rand_final_V)

df_TW_A <- rand_eob_A %>% 
  left_join(conv_A, by = "name") %>% 
  left_join(rand_final_A, by = "name") %>% 
  pivot_longer(c(rand_eob_diff, conv_diff, rand_final_diff), names_to = "method", values_to = "diff")

>>>>

---
title: "Supplementary material"

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72

@@

# knitr setup
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F,
                      dev = "jpeg",
                      dpi = 300,
                      fig.show = "hold",
                      fig.pos = "b", 
                      fig.path = "../figs/supplementary/figs/")

@@

```{r absnull, fig.cap="Comparison between the two M&V methods in detecting no intervention effect (expressed without absolute calculation) when buildings are subject to baseline change.", fig.height=6, fig.width=12}
#### ABS-NULL ####
rand_eob_S <- df_seq_fs_null_stable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = abs(0 - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_S <- df_fs_null_stable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_S <- df_fs_null_stable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

@@

# plot for the variable subset
rand_eob_V <- df_seq_fs_null_variable %>% 
  filter(seq == "eob") %>% 
  mutate(diff = abs(0 - fs), 
         method = "rand_eob") %>% 
  select(name, method, diff)

conv_V <- df_fs_null_variable %>% 
  filter(method != "rand") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - conv), 
         method = "conv") %>% 
  select(name, method, diff)

rand_final_V <- df_fs_null_variable %>% 
  filter(method != "conv") %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  mutate(diff = abs(true - rand), 
         method = "rand_final") %>% 
  select(name, method, diff)

@@

df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute effect in fractional savings", 
       title = "Absolute effect of non-routine events when using different M&V methods", 
       subtitle = str_glue("All {A_building} buildings wither measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))
```

```{r meancont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy (expressed without absolute calculation for errors)", fig.height=8, fig.width=12}

#### MEAN-CONT ####
df_MW_S <- df_cont_stable %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - cont_fs, 
         diff_all = savings - cont_fs_all) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_stable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = true - rand, 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))

# plot for the variable subset
df_MW_V <- df_cont_variable %>% 
  left_join(df_fs_variable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - cont_fs, 
         diff_all = savings - cont_fs_all) %>% 
  select(name, diff, diff_all, ratio) %>% 
  pivot_longer(c(diff, diff_all), names_to = "type", values_to = "values") %>% 
  mutate(ratio = ifelse(ratio == 1 & type == "diff", "1_1", 
                        ifelse(ratio == 1 & type == "diff_all", "1_2", 
                               ifelse(ratio == 2 & type == "diff", "2_1", 
                                      ifelse(ratio == 2 & type == "diff_all", "2_2", ratio))))) %>% 
  bind_rows(df_fs_variable %>% 
              filter(method != "conv") %>% 
              pivot_wider(names_from = method, values_from = savings) %>% 
              mutate(values = true - rand, 
                     ratio = "0", 
                     type = "diff_ref") %>% 
              select(name, values, ratio, type))
